Between Subgraph Isomorphism and Maximum Common Subgraph

### 2016-12-01 paper/paper.tex
Don't use microtype to avoid inevitable whining
New improved page layout
Little tweaks

\usepackage{times}
\usepackage{helvet}
\usepackage{courier}
\usepackage{complexity}
\usepackage{microtype}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{cleveref}
\usepackage{tikz}
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage[vlined]{algorithm2e}

@@

\begin{figure}[b]
    \centering
    \begin{tikzpicture}[scale=0.5]%[every node/.style={circle, inner sep=0.5mm,node
        %distance=10mm,fill=black}, scale=0.2]

@@

In this work we discuss an intermediate problem, where we must map all but $k$ vertices of the
pattern graph into the target---we show an example in \cref{figure:subgraphexample}. This in some
ways resembles the approximate subgraph matching model of \citet{DBLP:conf/cp/ZampelliDD05},
although we allow any vertex to be removed. We show that if
$k$ is reasonably small (say, between 1 and 5), then weakened forms of degree and path based
filterings are still effective in pruning the initial search space and in providing additional
constraints respectively. We then show that combining these techniques leads to a practical
algorithm which can scale to work with the families of graphs commonly used to benchmark subgraph
isomorphism algorithms: depending upon the benchmark family, we can close a substantial portion of
the instances, and in many more cases, we can at least obtain an upper bound. This is a significant
improvement over state of the art maximum common subgraph solvers, which cannot even fit many of
these instances in 64 GBytes of RAM. Finally, we show that iteratively increasing $k$ gives a
competitive algorithm for the maximum common induced subgraph problem.

@@

Throughout, our graphs are undirected, but may have loops.  We write $\V(G)$ for the vertex set of a
graph $G$, and $\EdgeSet(G)$ for its set of edges. We write $\N_G(v)$ for the \emph{neighbourhood}
of a vertex $v$ in $G$, and $v \sim_G w$ to mean $v$ is adjacent to $w$; where the graph is clear
from the context, the subscript is omitted.

A \emph{non-induced subgraph isomorphism} from a graph $P$ to a graph $T$ is an injective mapping
$\V(P) \rightarrowtail \V(T) $ which maps adjacent vertices to adjacent vertices. An \emph{induced subgraph
isomorphism} $P \hookrightarrow T$ additionally maps non-adjacent vertices to non-adjacent vertices.
We define a \emph{$k$-less subgraph isomorphism} from $P$ to $T$ to be a subgraph isomorphism from all but
$k$ vertices of $P$ to $T$; this may be non-induced or induced, written $P \lessnonind{k} T$ and $P
\lessind{k} T$ respectively. We write $p \lessmap{k} t$ to mean that the pattern vertex p is mapped
to the target vertex v under either kind of mapping.

@@

\begin{proposition}
    \label{prop:comp}
Let $i$ be an assignment of vertices of $T$ to vertices of $P$.  Then $i$ satisfies the definition of $P
\hookrightarrow T$ if and only if $i$ satisfies $P \rightarrowtail T$ and
$\loopcomp{P} \rightarrowtail \loopcomp{T}$.
Similarly, $i$ satisfies the definition of $P
\lessind{k} T$ if and only if $i$ satisfies both $P \lessnonind{k} T$ and
$\loopcomp{P}
\lessnonind{k} \loopcomp{T}$.
\end{proposition}

@@

Constraints are used to ensure that adjacent pairs of vertices are mapped to adjacenct pairs of
vertices. For the maximum common subgraph case, any pair of assignments with $\bot$ as either value
is also permitted.  To handle the induced case, rather than the usual approach of directly
having constraints for non-adjacent pairs of vertices, we will make use of \cref{prop:comp} by
seeking a mapping $i$ which simultaneously satisfies $P \rightarrowtail T$ and $\loopcomp{P}
\rightarrowtail \loopcomp{T}$. Finally, an ``all different except $\bot$ constraint'' ensures
injectivity.

@@

We perform our experiments on systems with dual Intel Xeon CPU E5-2640 processors with 64GBytes of
RAM, running Ubuntu 14.04. Our algorithm was implemented in
C++\footnote{https://github.com/ciaranm/aaai17-between-subgraph-isomorphism-and-maximum-common-subgraph-paper},
and source code was compiled using GCC 5.3.0. We used a timeout of 1,000 seconds for each instance.
For comparison, we used the implementations of \citet{DBLP:conf/cp/McCreeshNPS16}.

For datasets, we use the same 5,725 instances used in a recent work on portfolios of algorithms for
subgraph isomorphism \citep{thelionpaper}. This collection includes randomly generated scale-free
graphs \citep{constraints10}, an assortment of real-world graphs of varying sizes \citep{LV02},
randomly generated graph pairs (using bounded degree, regular mesh, and uniform models; all are
satisfiable), segmented images~\citep{pr15,cviu11}, meshes from modelling 3D objects \citep{cviu11},
and graphs close to the satisfiable-unsatisfiable phase transition
\citep{DBLP:conf/ijcai/McCreeshPT16}. All of these instances are available in a simple text
format\footnote{http://liris.cnrs.fr/csolnon/SIP.html}.  Note that many of these instances
are \emph{much} larger than were used in the recent comparison of maximum common subgraph algorithms
by \citet{DBLP:conf/cp/McCreeshNPS16}: for unlabelled graphs, they used pairs of graphs with up to
35 vertices in each, whilst our dataset contains graphs with up to 6,671 vertices. For the
forward checking algorithm described by \citet{DBLP:conf/cp/NdiayeS11}, 1,560 of the instances
cannot fit in the amount of RAM we have available---we treat these instances as having timed out.
The situation for the clique encoding is even worse, and 3,653 of these instances do not fit in
64GBytes of RAM. In contrast, for our algorithm, every instance fits comfortably.

@@

Let $p$ be a vertex in a graph $P$. The \emph{degree} of $p$, $\deg(p)$, is the number of vertices
to which it is adjacent. The degree of vertices gives us an invariant, which may be used to
eliminate some infeasible values from domains.

@@

The \emph{neighbourhood degree sequence} (NDS) of a vertex $p$, $\nds(p)$, is the (non-ascending)
sequence of degrees of its neighbours.

@@

\begin{algorithm*}[tb]

@@

@@

We begin by trying to infer domain deletions. The $\FuncSty{propagate}$ function looks for domains
which contains either only a single value, or only wildcards---we call such a domain
\emph{effectively-unit}. If such a domain $D_v$ exists, we eliminate its value from every other domain,
and then propagate adjacency: for each domain corresponding to a vertex adjacent to $v$, we
eliminate any value from its domain which is not adjacent to $v$ (treating wildcards as being
adjacent to all vertices). If a domain wipeout occurs, we return failure.

@@

To handle the induced case, we make use of \cref{prop:comp} (line~\ref{line:comp}). We also tested
using path reasoning on complement graphs, but this is expensive to calculate and provides little
benefit in practice on the instances we considered.  We also do not strip isolated vertices as the
original algorithm did, as this is not a valid simplification in the induced case, and we do not use
conflict-directed backjumping as in our experiments it had very little effect for $k > 0$.

@@

\begin{figure*}[tb]
    \centering
    \includegraphics*{gen-graph-runtimes.pdf}
    \caption{In the first two plots we show the cumulative number of instances solved over time, for the
        induced and non-induced problems, with different values of $k$. We also show the results of
        iteratively increasing $k$ until a solution is found, and in the induced case, the
        performance of two leading maximum common subgraph algorithms. In the third plot we show results
        comparing iteratively increasing $k$ with our algorithm to other approaches on maximum
        common induced subgraph instances.}\label{figure:runtimes}
\end{figure*}

\begin{figure}[ht!] % This is necessary to avoid really awful figure placement midway through the
    % references...
    \centering
    \includegraphics*{gen-graph-which-k-by-family.pdf}
    \caption{The proportion of instances, in different families, which become satisfiable for
    increasing values of $k$. The ``larger'' instances are those where we can prove unsatisfiability for
    $k = 5$, whilst the gap between the top of the bar and the top of the graph is the fraction of
    instances where a timeout was reached for at least one value of $k$.}\label{figure:which-k}
\end{figure}

>>>>

\usepackage{times}
\usepackage{helvet}
\usepackage{courier}
\usepackage{complexity}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{cleveref}
\usepackage{tikz}
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage[vlined]{algorithm2e}

@@

\begin{figure}[b]
    \centering
    \begin{tikzpicture}[scale=0.58]%[every node/.style={circle, inner sep=0.5mm,node
        %distance=10mm,fill=black}, scale=0.2]

@@

In this work we discuss an intermediate problem, where we must map all but $k$ vertices of the
pattern graph into the target---we show an example in \cref{figure:subgraphexample}. This in some
ways resembles the approximate subgraph matching model of \citet{DBLP:conf/cp/ZampelliDD05},
although we allow any vertex to be removed. We show that if
$k$ is reasonably small (say, between 1 and 5), then weakened forms of degree and path based
filterings are still effective in pruning the initial search space and in providing additional
constraints respectively. We then show that combining these techniques leads to a practical
algorithm which can scale to work with the families of graphs commonly used to benchmark subgraph
isomorphism algorithms: depending upon the benchmark family, we can close a substantial portion of
the instances, and in many more cases, we can at least obtain an upper bound. This is a significant
improvement over state of the art maximum common subgraph solvers, which cannot even fit many of
these instances in 64 GBytes of RAM. Finally, we show that starting with $k = 0$ and iteratively
increasing $k$ gives a competitive algorithm for the maximum common induced subgraph problem.

@@

Throughout, our graphs are undirected, but may have loops.  We write $\V(G)$ for the vertex set of a
graph $G$, and $\EdgeSet(G)$ for its set of edges. We write $\N_G(v)$ for the \emph{neighbourhood}
of a vertex $v$ in $G$, and $v \sim_G w$ to mean $v$ is adjacent to $w$. The \emph{degree} of $p$,
$\deg_G(p)$, is the number of vertices to which it is adjacent. Where the graph is clear from the
context, the subscripts are omitted.

A \emph{non-induced subgraph isomorphism} from a graph $P$ (called the \emph{pattern}) to a graph
$T$ (the \emph{target}) is an injective mapping $\V(P) \rightarrowtail \V(T) $ which maps adjacent
vertices to adjacent vertices. An \emph{induced subgraph isomorphism} $P \hookrightarrow T$
additionally maps non-adjacent vertices to non-adjacent vertices.  We define a \emph{$k$-less
subgraph isomorphism} from $P$ to $T$ to be a subgraph isomorphism from all but $k$ vertices of $P$
to $T$; this may be non-induced or induced, written $P \lessnonind{k} T$ and $P \lessind{k} T$
respectively. We write $p \lessmap{k} t$ to mean that the pattern vertex p is mapped to the target
vertex v under either kind of mapping.

@@

\begin{proposition}
    \label{prop:comp}
Let $i$ be an assignment of vertices of $T$ to vertices of $P$.  Then $i$ satisfies the definition of $P
\hookrightarrow T$ if and only if $i$ satisfies $P \rightarrowtail T$ and
$\loopcomp{P} \rightarrowtail \loopcomp{T}$ simultaneously.
Similarly, $i$ satisfies the definition of $P
\lessind{k} T$ if and only if $i$ satisfies both $P \lessnonind{k} T$ and
$\loopcomp{P}
\lessnonind{k} \loopcomp{T}$.
\end{proposition}

@@

Constraints are used to ensure that adjacent pairs of vertices are mapped to adjacenct pairs of
vertices. For the maximum common subgraph case, any pair of assignments with $\bot$ as either value
is also permitted.  To handle the induced case, rather than the usual approach of directly
having constraints for non-adjacent pairs of vertices, we will be making use of \cref{prop:comp} by
seeking a mapping $i$ which simultaneously satisfies $P \rightarrowtail T$ and $\loopcomp{P}
\rightarrowtail \loopcomp{T}$. Finally, an ``all different except $\bot$ constraint'' ensures
injectivity.

@@

We perform our experiments on systems with dual Intel Xeon CPU E5-2640 processors with 64GBytes of
RAM, running Ubuntu 14.04. Our algorithm was implemented in
C++\footnote{https://github.com/ciaranm/aaai17-between-subgraph-isomorphism-and-maximum-common-subgraph-paper},
and source code was compiled using GCC 5.3.0. We used a timeout of 1,000 seconds for each instance.
For comparison, we used the implementations of \citet{DBLP:conf/cp/NdiayeS11} and \citet{DBLP:conf/cp/McCreeshNPS16}.

For datasets, we use the same 5,725 instances used in a recent work on portfolios of algorithms for
subgraph isomorphism \citep{thelionpaper}. This collection includes randomly generated scale-free
graphs \citep{constraints10}, an assortment of real-world graphs of varying sizes \citep{LV02},
randomly generated graph pairs (using bounded degree, regular mesh, and uniform models; all are
satisfiable), segmented images~\citep{pr15,cviu11}, meshes from modelling 3D objects \citep{cviu11},
and graphs close to the satisfiable-unsatisfiable phase transition
\citep{DBLP:conf/ijcai/McCreeshPT16}. All of these instances are available in a simple text
format\footnote{http://liris.cnrs.fr/csolnon/SIP.html}.  Note that many of these instances
are \emph{much} larger than were used in the recent comparison of maximum common subgraph algorithms
by \citet{DBLP:conf/cp/McCreeshNPS16}: for unlabelled graphs, they used pairs of graphs with up to
35 vertices in each, whilst our dataset contains graphs with up to 6,671 vertices. For the
forward-checking algorithm described by \citet{DBLP:conf/cp/NdiayeS11}, 1,560 of the instances
cannot fit in the amount of RAM we have available---we treat these instances as having timed out.
The situation for the clique encoding is even worse, and 3,653 of these instances do not fit in
64GBytes of RAM. In contrast, for our algorithm, every instance fits comfortably.

@@

Let $p$ be a vertex in a graph $P$. The degree of vertices gives us an invariant, which may be used to
eliminate some infeasible values from domains as follows.

@@

The \emph{neighbourhood degree sequence} (NDS) of a vertex $p$, $\nds(p)$, is the (non-ascending)
sequence of degrees of its neighbours. \citet{constraints10} showed how this may be used for
filtering in subgraph isomorphism. We extend this for the $k$-less setting as follows.

@@

\begin{algorithm*}[p]

@@

\begin{figure*}[p]
    \centering
    \includegraphics*{gen-graph-runtimes.pdf}
    \caption{In the first two plots we show the cumulative number of instances solved over time, for the
        induced and non-induced problems, with different values of $k$. We also show the results of
        iteratively increasing $k$ until a solution is found, and in the induced case, the
        performance of two leading maximum common subgraph algorithms. In the third plot we show results
        comparing iteratively increasing $k$ with our algorithm to other approaches on maximum
        common induced subgraph instances.}\label{figure:runtimes}
\end{figure*}

@@

We begin by trying to infer domain deletions. The $\FuncSty{propagate}$ function looks for domains
which contain either only a single value, or only wildcards---we call such a domain
\emph{effectively-unit}. If such a domain $D_v$ exists, we eliminate its value from every other domain,
and then propagate adjacency: for each domain corresponding to a vertex adjacent to $v$, we
eliminate any value from its domain which is not adjacent to $v$ (treating wildcards as being
adjacent to all vertices). If a domain wipeout occurs, we return failure.

@@

To handle the induced case, we make use of \cref{prop:comp} (line~\ref{line:comp}). We considered
using path reasoning on complement graphs, but this is expensive to calculate and provides little
benefit in practice on these instances.  We also do not strip isolated vertices as the
original algorithm did, as this is not a valid simplification in the induced case, and we do not use
conflict-directed backjumping as in our experiments it had very little effect for $k > 0$.

@@

\begin{figure}[t]
    \centering
    \includegraphics*{gen-graph-which-k-by-family.pdf}
    \caption{The proportion of instances, in different families, which become satisfiable for
    increasing values of $k$. The ``larger'' instances are those where we can prove unsatisfiability for
    $k = 5$, whilst the gap between the top of the bar and the top of the graph is the fraction of
    instances where a timeout was reached for at least one value of $k$.}\label{figure:which-k}
\end{figure}
